### 2021-05-05 Paper.tex
modified

\documentclass[hidelinks,11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[cm]{fullpage}
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{doi}
\usepackage{bm}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
}
\usepackage{float}
\restylefloat{table}
\usepackage{siunitx}
\DeclareSIUnit\year{yr}

@@

\begin{table}[H]
\caption{List of benchmarks implemented for each feature of the code.}
\centering
\begin{tabular}{| l | l |}
\hline
 \textbf{Feature}  & \textbf{Benchmark}  \\
\hline
\hline
  Solver  & Stokes flow   \\
\hline
  4th-order Runge-Kutta  &  \\
  markers advection  & Zalesak disk   \\
\hline
  Momentum equation  & Poiseuille flow   \\
    & Instantaneous 2D sphere   \\
    & Rayleigh-Taylor experiment  \\
    & Falling block  \\
\hline
  Sticky air and  & 2D Stokes sphere \\
  free surface  & Stabilisation algorithm   \\
  & Topography relaxation   \\
  & Spontaneous subduction   \\
\hline
  Erosion and sedimentation  & -  \\
\hline
   Non-linear  & Slab detachment   \\
   visco-plasticity & Indenter   \\
    & Brick   \\
\hline
  Energy equation  & Advection stabilisation   \\
    & Simple shear heating \\
    & Shear and adiabatic heating \\
    & Mantle convection   \\
    & Thin layer entrainment   \\
\hline
  Phase changes  &  \\
  and hydration  & Hydrated sinking cylinder   \\
\hline
  Melting & Experimental melting curves \\
\hline
\end{tabular}
\label{tab:list}
\end{table}

@@

Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average
\citep{Tackley2010,Glerum2018} as follows
\begin{equation}\label{eq:viscoplastic}
\eta_{vp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}+\frac{1}{\eta_{pl}}\right)^{-1}
\end{equation}
Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
low or high viscosity \citep{Glerum2018} as follows
\begin{equation}\label{eq:effective}
\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
\end{equation}
Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
\begin{equation}\label{eq:convergence}
\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
\end{equation}
%\begin{equation}\label{eq:convergence}
%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
%\end{equation}
%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
$\bm{\mathcal{R}}^i$ and it is normalised as follows
\begin{equation}\label{eq:convergence}
\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
\end{equation}
where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of the velocity residual is calculated as
\begin{equation}\label{eq:convergence_vel}
\frac{||\bm{v}_1-\bm{v}_{1-1}||_2}{||\bm{v}_{i-1}||_2}
\end{equation}
Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or both the normalised nonlinear residual and the
velocity residual converge under a defined tolerance $tol$.

@@

\section{Benchmarks}\label{sec:benchmark}
\subsection{Error measurements}\label{sec:error}
In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
\begin{eqnarray}
err_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
err_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
\end{eqnarray}
respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.

Other important quantities calculated in some benchmarks are the root-mean-square velocity and the Nusselt number, calculated as
\[v_{rms}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
and
\[Nu=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
respectively.

@@

\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
We distinguish three different types of velocity boundary conditions:
\begin{enumerate}
\item FS: free slip conditions on all sides.
\item NS: no slip conditions on all sides.
\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
\end{enumerate}
All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{rms}$ and average pressure on the
entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{rms}$ for NS model are also shown
in Fig. \ref{fig:inst_sphere}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
\caption{$v_{rms}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
\label{fig:inst_sphere}
\end{figure}

@@

$v_{rms}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
\caption{$v_{rms}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
\label{fig:RT}
\end{figure}

@@

\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{rms}$ in case of an arithmetic average are also
shown in Fig. \ref{fig:inst_spherefs}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
\caption{$v_{rms}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
Results are compared with results obtained by ASPECT.}
\label{fig:inst_spherefs}
\end{figure}

@@

\subsection{Mantle convection}\label{sec:mantle}
This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
resolution (between \num{32x32} and \num{128x128} elements). Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity
($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
\[T(x,y)=(1-y)-0.01 \cos (\pi x) \sin (\pi x)\]

The solution generated by the code in terms of $v_{rms}$ (calculated as in Sec. \ref{sec:error}) and $Nu$ as function of time are reported for all the
simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{rms}$ and $Nu$ of all simulations converge well toward the values from \citet{Blankenbach1989},
with lower errors for higher resolution grids (Table \ref{tab:mantle}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
\caption{$v_{rms}$ (panels a, c and e) and $Nu$ (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
indicate the convergence values for the $v_{rms}$ and the $Nu$ from \citet{Blankenbach1989} at the steady state.}
\label{fig:mantle}
\end{figure}

\begin{table}[H]
\caption{Comparison between $v_{rms}$ and $Nu$ predicted by the code for the mantle convection experiment and same values as reported in literature.}
\centering
\small
\begin{tabular}{|p{0.09\textwidth}|p{0.05\textwidth}|p{0.24\textwidth}|p{0.15\textwidth}|p{0.1\textwidth}p{0.1\textwidth}p{0.1\textwidth}|}
\hline
  & & \citet{Blankenbach1989} & \citet{Thieulot2014} & & &   \\
 & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
\hline
\hline
 $Ra=10^4$ & $v_{rms}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
           & $Nu$      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
\hline
 $Ra=10^5$ & $v_{rms}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
           & $Nu$      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
\hline
 $Ra=10^6$ & $v_{rms}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
           & $Nu$      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
\hline
\end{tabular}
\label{tab:mantle}
\end{table}

@@

The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{rms}$
(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
\caption{$v_{rms}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
\label{fig:thin}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
\caption{$v_{rms}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
\label{fig:thin_initial}
\end{figure}

\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
This benchmark is performed as described in \citet{Tosi2015}.

>>>>

\documentclass[hidelinks,10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[cm]{fullpage}
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{doi}
\usepackage{bm}
\usepackage[table]{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={blue},
    citecolor={blue},
    urlcolor={blue}
}
\usepackage{float}
\restylefloat{table}
\usepackage{siunitx}
\DeclareSIUnit\year{yr}

@@

\begin{table}[H]
\caption{List of benchmarks implemented for each feature of the code.}
\centering
\begin{tabular}{| l | l |}
\toprule
 \textbf{Feature}  & \textbf{Benchmark}  \\
\midrule
  Solver  & Stokes flow   \\
\hline
  4th-order Runge-Kutta  & \multirow{2}{*}{Zalesak disk} \\
  markers advection  &  \\
\hline
  \multirow{4}{*}{Momentum equation}  & Poiseuille flow   \\
    & Instantaneous 2D sphere   \\
    & Rayleigh-Taylor experiment  \\
    & Falling block  \\
\hline
    & 2D Stokes sphere \\
    Sticky air and  & Stabilisation algorithm   \\
    free surface & Topography relaxation   \\
  & Spontaneous subduction   \\
\hline
  Erosion and sedimentation  & -  \\
\hline
   Non-linear  & Slab detachment   \\
   visco-plasticity & Indenter   \\
    & Brick   \\
\hline
  \multirow{5}{*}{Energy equation}  & Advection stabilisation   \\
    & Simple shear heating \\
    & Shear and adiabatic heating \\
    & Mantle convection   \\
    & Thin layer entrainment   \\
\hline
  Phase changes  & \multirow{2}{*}{Hydrated sinking cylinder} \\
  and hydration  &  \\
\hline
  Melting & Experimental melting curves \\
\bottomrule
\end{tabular}
\label{tab:list}
\end{table}

@@

Viscous creep and plastic yielding are then combined to obtain a viscoplastic viscosity $\eta_{vp}$, which is computed with an harmonic average
\citep{Tackley2010,Glerum2018} as follows
\begin{equation}\label{eq:viscoplastic}
\eta_{vp}=\left(\frac{1}{\eta_{df}}+\frac{1}{\eta_{ds}}+\frac{1}{\eta_{pl}}\right)^{-1}
\end{equation}
Finally, effective viscosity $\eta_{eff}$ is capped by the minimum and the maximum viscosity ($\eta_{min}$ and $\eta_{max}$, respectively) to avoid extremely
low or high viscosity \citep{Glerum2018} as follows
\begin{equation}\label{eq:effective}
\eta_{eff}=\min(\max(\eta_{vp},\eta_{min}),\eta_{max})
\end{equation}
Viscous creep and plastic yielding are non-linear rheologies because of their dependence on the velocity field through pressure and strain rates. Therefore,
the solution is determined by means of Picard-type iterations, until convergence of the velocity field \citep{Glerum2018}. The convergence is verified at each
iteration $i$ via the nonlinear residual $\bm{\mathcal{R}}^i$ that can be determined as
\begin{equation}\label{eq:convergence}
\bm{\mathcal{R}}^i= \mathbb{K} (\eta_{eff}(\dot{\bm{\epsilon}}^{i-1},p^{i-1}))\cdot \bm{v}^{i-1} - \bm{f}^i
\end{equation}
%\begin{equation}\label{eq:convergence}
%\chi=1-\frac{\langle (\bm{v}^i-\langle\bm{v}^i\rangle)\cdot (\bm{v}^{i+1}-\langle\bm{v}^{i+1}\rangle) \rangle}{\sqrt{\sigma^i\sigma^{i+1}}}
%\end{equation}
%Iterations are performed until either convergence ($\chi<tol$) or maximal number of non-linear iterations ($it_{max}$) is reached.
where $\mathbb{K}$ is the Stiffness matrix and $\bm{f}$ is the right hand side vector \citep{Spiegelman2016,Glerum2018}.  The $L_2$-norm is extracted from
$\bm{\mathcal{R}}^i$ and it is normalised as follows
\begin{equation}\label{eq:convergence_res}
\frac{||\bm{\mathcal{R}}^i||_2}{||\bm{\mathcal{R}}^0||_2}
\end{equation}
where $||\bm{\mathcal{R}}^0||_2$ is the $L_2$-norm of the first nonlinear iteration.
Since often the normalised nonlinear residual drop very quickly over the first 2-3 iterations, also the $L_2$-norm of the velocity residual is calculated as
\begin{equation}\label{eq:convergence_vel}
\frac{||\bm{v}_1-\bm{v}_{1-1}||_2}{||\bm{v}_{i-1}||_2}
\end{equation}
Iterations are performed until either maximal number of non-linear iterations ($it_{max}$) is reached or both the normalised nonlinear residual and the
velocity residual converge under a defined tolerance $tol$.

@@

\section{Benchmarks}\label{sec:benchmark}
\subsection{Error measurements}\label{sec:error}
In order to determine the accuracy of velocity and pressure field of the benchmarks, the $L_2$-norm is computed by numerical integration on the Gauss-Legendre
quadrature points. $L_2$-norm for pressure and velocity errors can be evaluated as
\begin{eqnarray}
\textrm{err}_p&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|p^n_i(r_q)-p^a_i(r_q)|^2\right)w_q|J_q|}\nonumber \\
\textrm{err}_v&=&\sqrt{\sum_{i=1}^{n_e}\sum_{q=1}^{n_q}\left(|u^n_i(r_q)-u^a_i(r_q)|^2+|v^n_i(r_q)-v^a_i(r_q)|^2\right)w_q|J_q|}\nonumber
\end{eqnarray}
respectively, where $n_e$ is the number of elements, $n_q$ is the number of quadrature points per element, $p^n_i(r_q)$ and $p^a_i(r_q)$ are the numerical and
analytical pressure, respectively, in each quadrature point $q$, $w_q$  and $J_q$ are the weight and the Jacobian at the quadrature point
$q$, $u^n_i(r_q)$, $v^n_i(r_q)$, $u^a_i(r_q)$ and $v^a_i(r_q)$ are the numerical and analytical velocities, respectively, in each quadrature point $q$.

Other important quantities calculated in some benchmarks are the root-mean-square velocity and the Nusselt number, calculated as
\[v_{\textrm{rms}}=\sqrt{\frac{1}{V_{\Omega}}\int_{\Omega}|\bm{v}|^2 d\Omega}\]
and
\[\textrm{Nu}=\frac{\int_{top}\bm{\nabla} T \cdot \bm{n} ds}{\int_{bottom}T ds}\]
respectively.

@@

\subsection{Instantaneous 2D sphere}\label{sec:ist_sphere}
The domain is a unit square with gravity $\bm{g}=(0,-1)$. The fluid has constant density and viscosity ($\rho_f=1$ and $\eta_f=1$). The sphere is in the middle
of the domain with a radius R = 0.123456798 and has constant density and viscosity ($\rho_s=10^{-2}+\rho_f$ and $\eta_s=10^3 \cdot \eta_f$).
We distinguish three different types of velocity boundary conditions:
\begin{enumerate}
\item FS: free slip conditions on all sides.
\item NS: no slip conditions on all sides.
\item OT: free slip conditions on the sides and the bottom, and open boundary at the top.
\end{enumerate}
All three cases are performed for different grid resolutions, between \num{16x16} and \num{512x512} elements with 50 randomly distributed markers per element,
and different average schemes for the viscosity (harmonic, geometric and arithmetic). The velocity in the centre of the sphere $\bm{v}(0.5;0.5)$), minimum and
maximum velocities ($u_{min}$, $u_{max}$, $v_{min}$, $v_{max}$) and pressures ($p_{min}$, $p_{max}$) on the entire domain, $v_{\textrm{rms}}$ and average pressure on the
entire domain ($p_{avg}$) are compared with solutions generated by ASPECT \citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and all results
can be found at \url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere2D}. Results in terms of the $v_{\textrm{rms}}$ for NS model are also shown
in Fig. \ref{fig:inst_sphere}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_NS.pdf}
\caption{$v_{\textrm{rms}}$ as function of element size for the instantaneous 2D sphere experiment in case of no slip boundary conditions and with different average
schemes for the viscosity. Results are compared with results obtained by other numerical codes.}
\label{fig:inst_sphere}
\end{figure}

@@

$v_{\textrm{rms}}$ as function of time is reported in Fig. \ref{fig:RT}, matching well with results shown by \citet{vanKeken1997}, \citet{Tackley2003} and
\citet{Thieulot2014}. Fig. \ref{fig:rayleigh} shows the evolution of the experiment at different time steps.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/RT.pdf}
\caption{$v_{\textrm{rms}}$ of the Rayleigh-Taylor experiment as function of time for different resolution of the grid.}
\label{fig:RT}
\end{figure}

@@

\subsection{2D time-dependent Stokes sphere with deformable free surface}\label{sec:sphere}
This experiment is performed in a unit square domain with the gravity acceleration fixed to $g_y=-1$. The sphere has $\rho_s=2$ and $\eta_s=10^3$ and is
initially centred at ($0.5;0.6$) with radius $R=0.123456789$. The fluid surrounding the sphere has $\rho_f=1$ and $\eta_f=1$ and occupies the domain for
$y\leq 0.75$, while for $y>0.75$ the air has $\rho_a=0$ and $\eta_a=10^{-3}$. Velocity boundary conditions are set to free slip on all sides. The Courant
number is set to 0.25. The experiments run for 200 s using grid resolutions from \num{150x150} to \num{512x512} elements, with 25 randomly distributed markers
per element, and for different average schemes for the viscosity. The interface between the fluid and the air is tracked by means of the markers chain. Results
in terms of velocity and pressure fields and topography variations are compared with results obtained with ASPECT
\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and they can be found at
\url{https://github.com/cedrict/fieldstone/tree/master/images/stokes_sphere_fs2D}. Results in terms of the $v_{\textrm{rms}}$ in case of an arithmetic average are also
shown in Fig. \ref{fig:inst_spherefs}.

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_arithm.pdf}
\caption{$v_{\textrm{rms}}$ for different grid resolutions as function of time for the 2D time-dependent Stokes sphere experiment in case of an arithmetic average.
Results are compared with results obtained by ASPECT.}
\label{fig:inst_spherefs}
\end{figure}

@@

\subsection{Mantle convection}\label{sec:mantle}
This problem is performed as presented by \citet{Blankenbach1989} (constant viscosity cases) and \citet{Thieulot2014}, in a 2D unit square domain with gravity
acceleration $g_y=10^{10}Ra$. The experiment is performed with three different Rayleigh numbers ($Ra=10^4$, $10^5$ and $10^6$) and with different grid
resolution (between \num{32x32} and \num{128x128} elements). Both fluids have constant viscosity, initial density, heat capacity, thermal conductivity
($\eta= \rho_0= C_p = k=1$), reference temperature ($T_0=0$) and thermal expansion coefficient ($\alpha=10^{-10}$). Temperatures are set to 0 on top and 1 on
bottom of the domain. Velocity boundary conditions are set to free slip on all sides. The initial temperature field is given by
\[T(x,y)=(1-y)+0.01 \cos (\pi x) \sin (\pi y)\]

The solution generated by the code in terms of $v_{\textrm{rms}}$ (calculated as in Sec. \ref{sec:error}) and Nu as function of time are reported for all the
simulations in Fig. \ref{fig:mantle}. At the steady state, $v_{\textrm{rms}}$ and Nu of all simulations converge well toward the values from \citet{Blankenbach1989},
with lower errors for higher resolution grids (Table \ref{tab:mantle}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/Convection.pdf}
\caption{$v_{\textrm{rms}}$ (panels a, c and e) and Nu (panels b, d and f) for the mantle convection benchmark as function of time for different grid resolution.
Panels a and b show the results for $Ra=10^4$; Panels c and d show the results for $Ra=10^5$; panels e and f show the results for $Ra=10^6$. Purple lines
indicate the convergence values for the $v_{\textrm{rms}}$ and the Nu from \citet{Blankenbach1989} at the steady state.}
\label{fig:mantle}
\end{figure}

\begin{table}[H]
\caption{Comparison between $v_{\textrm{rms}}$ and Nu predicted by the code for the mantle convection experiment and same values as reported in literature.}
\centering
\small
\begin{tabular}{|>{\centering}m{0.09\textwidth}|>{\centering}m{0.05\textwidth}|>{\centering}m{0.24\textwidth}|>{\centering}m{0.15\textwidth}|
  >{\centering}m{0.1\textwidth} >{\centering}m{0.1\textwidth} >{\centering\arraybackslash}m{0.1\textwidth}|}
\toprule
& & \multirow{2}{*}{\citet{Blankenbach1989}} & \citet{Thieulot2014} & \multicolumn{3}{c|}{FALCON}  \\
 & & & \num{200x200} & \num{32x32} & \num{64x64} & \num{128x128}  \\
\midrule
\multirow{2}{*}{$Ra=10^4$} & $v_{\textrm{rms}}$ & $42.864947 \pm 0.000020$ & 42.867 & 42.83226 & 42.852793 & 42.861394  \\
           & Nu      & $4.884409 \pm 0.000010$  & 4.882  & 4.781297 & 4.857475  & 4.877573   \\
\hline
\multirow{2}{*}{$Ra=10^5$} & $v_{\textrm{rms}}$ & $193.21454 \pm 0.00010$  & 193.255& 193.872643 & 193.377472 & 193.252290  \\
           & Nu      & $10.534095 \pm 0.000010$ & 10.507 & 9.602514 & 10.270735  & 10.465629   \\
\hline
\multirow{2}{*}{$Ra=10^6$} & $v_{\textrm{rms}}$ & $833.98977 \pm 0.00020$  & 834.712& 848.091176 & 837.767911 & 834.945793  \\
           & Nu      & $21.972465 \pm 0.000020$ & 21.695 & 15.999266 & 19.703682  & 21.306939   \\
\bottomrule
\end{tabular}
\label{tab:mantle}
\end{table}

@@

The experiment is performed with two grid resolution (\num{125x40} and \num{200x80} elements), with 100 markers per element and CFL of 0.25. $v_{\textrm{rms}}$
(see Sec. \ref{sec:error}) calculated for both the simulations match well with results from \citet{vanKeken1997} and obtained with ASPECT
\citep{KHB12,heister_aspect_methods2,aspect-doi-v2.2.0,aspectmanual} and ELEFANT \citep{Thieulot2014}, obtained with similar grid resolutions
(Fig. \ref{fig:thin}). In addition, a variety of simulations with different resolutions and aspect ratios are performed to verify that the sensitivity of the
initial velocity field. Results match well with results from \citet{vanKeken1997} (in the grey area) and \citet{Thieulot2014} (Fig. \ref{fig:thin_initial}).

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrms_thin.pdf}
\caption{$v_{\textrm{rms}}$ for the thin layer experiment as function of time for different grid resolution. Results are compared with results obtained by
\citet{vanKeken1997} (black lines) and with ASPECT and ELEFANT.}
\label{fig:thin}
\end{figure}

\begin{figure}
\centering
\noindent\includegraphics[width=400px]{./Figures/vrmszero.pdf}
\caption{$v_{\textrm{rms}}$ at $t=0$ for the thin layer experiment as function of the element size for different aspect ratios of the elements. Results are compared
with results obtained by \citet{vanKeken1997} (black dots in the grey area) and with ASPECT and ELEFANT.}
\label{fig:thin_initial}
\end{figure}

\subsection{Viscoplastic mantle convection}\label{sec:visco_mantle}
This benchmark is performed as described in \citet{Tosi2015} for Cases 1-5a. Table \ref{tab:visco_mantle}.WORK IN PROGRESS.

\renewcommand{\arraystretch}{1.5}
{\rowcolors{3}{white!100!}{gray!25!}
\begin{table}[H]
  \caption{Results of the viscoplastic mantle convection benchmark for Cases 1-4 compared with results obtained with ELEFANT and YACC (from \citealp{Tosi2015}).}
  \centering
  \resizebox{\textwidth}{!}{
\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
  \toprule
    & \multicolumn{3}{c|}{\textbf{Case 1}} & \multicolumn{3}{c|}{\textbf{Case 2}} & \multicolumn{3}{c|}{\textbf{Case 3}} &
    \multicolumn{3}{c|}{\textbf{Case 4}} \\
   & FALCON & ELEFANT& YACC & FALCON& ELEFANT& YACC & FALCON & ELEFANT & YACC & FALCON & ELEFANT & YACC \\
  \midrule
   $\left\langle T \right\rangle$ & 0.7759 & 0.7758 & 0.7767 & & 0.6036 & 0.6058 & 0.7275 & 0.7275 & 0.7286 & & 0.5277 & 0.5289\\
    Nu$^{top}$ & 3.4280 & 3.4214 & 3.4298 & & 8.5115 & 8.5278 & 3.0368 & 3.0347 & 3.0374 && 6.5912 & 6.5572 \\
    Nu$^{bottom}$ & 3.3168 & 3.313 & 3.3143 & & 8.4465 & 8.3990 & 3.0086 & 2.9908 & 2.9628 && 6.5834 & 6.5243 \\
    $v_{\textrm{rms}}$ & 249.901 & 249.134 & 251.7997 & & 140.7087 & 142.2020 & 100.2776 & 100.1208 & 100.9467 && 79.1105 & 79.6202 \\
    $v_{\textrm{rms}}^{top}$ & 1.8907 & 1.8642 & 1.8298 & & 104.1228 & 105.1767 & 2.0832 & 2.0652 & 2.0374 && 74.7596 & 75.8414 \\
    $u_{max}^{top}$ & 2.6340 & 2.6119 & 2.5516 & & 121.766 & 122.3238 & 2.9086 & 2.9019 & 2.8458 && 88.9146 & 89.2940 \\
    $\eta_{min}$ & - & - & - & & $2.0600\times10^{-5}$ & $1.9853\times10^{-5}$ & $4.8129\times10^{-5}$ & $4.8080\times10^{-5}$ & $4.7907\times10^{-5}$ &
    & $1.9860\times10^{-4}$ & $1.9174\times10^{-4}$ \\
    $\eta_{max}$ & - & - & - & & 1.85 & 1.6095 & 0.9360 & 0.9023 & 1 && 1.52 & 1.6773 \\
    $\left\langle W \right\rangle$ & 2.4299 & 2.4316 & 2.4583 & & 7.5665 & 7.6222 & 2.0387 & 2.0384 & 2.0400 && 5.6216 & 5.6512 \\
    $\left\langle \Phi \right\rangle/Ra$ & 2.4311 & 2.4276 & 2.4333 & & 7.5585 & 7.6194 & 2.0386 & 2.0370 & 2.0335 && 5.6182 & 5.6463 \\
    $\delta$ & 0.05\% & 0.16\% & 1.02\% & & 0.11\% & 0.04\% & $<$0.01\% & 0.07\% & 0.32\% && 0.06\% & 0.09\% \\
  \bottomrule
  \end{tabular}}
  \label{tab:visco_mantle}
  \end{table}
